Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-02 proof_overview.tex
Major update to Section 1 and 2

The key result in this paper is Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}. In particular, it is the key ingredient in the proofs of Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}. Therefore, in this section we first give a detailed overview of the structure for the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}. The proof itself, based on intermediate propositions, can be found in Section~\ref{ssec:proof_main_results}. Here we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proves of Theorem~\ref{thm:exact_expression_c_infty} and Theorem~\ref{thm:exact_expression_local_clustering} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}

Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic} based on the different propositions described below, plus the sections in which theses propositions are proved.

@@

\caption{Overview of the proof strategy for Theorem \ref{thm:asymptotic_local_clustering_hyperbolic}.}
\label{fig:overview_proof}
\end{figure}

The main idea behind the proof of Theorem \ref{thm:asymptotic_local_clustering_hyperbolic} is to couple the hyperbolic graph $G_{\H, n}(\alpha, \nu)$ with a subgraph of an infinite graph $G_\Pcal(\alpha,\nu)$ (see Section~\ref{ssec:infinite_model}). A similar approach is used in \cite{fountoulakis2018law} to analyze the size of the largest component in the hyperbolic model and in \cite{komjathy2018explosion} to analyze explosion times in a more general class of models.. 
The advantage of the infinite limit model is two-fold. First, it simplifies the actual computations, e.g. by replacing hyperbolic densities with exponential densities and the hyperbolic distance with a computationally easier expression. Secondly, it directly yields the limit as a result rather than having to determine this later from a long expression depending on $n$.

Very roughly, the proof falls into three parts: justifying the transitioning from the hyperbolic random graph to the infinite limit model, proving concentration results for local clustering and calculating the expected local clustering coefficient and function in the infinite limit model.

In this section, we state the required definitions, auxiliary lemmas and propositions and show how to assemble them to obtain Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}.

To analyze the limiting expressions for clustering in the hyperbolic random graph $G_{\H,n}(\alpha, \nu)$ we shall use an following infinite graph model which was also considered in~\cite{fountoulakis2018law} to analyze the largest component. Define $\Rcal = \mathbb{R} \times \mathbb{R}_+$ and let $\mathcal{P}_{\alpha,\nu}$ be a Poisson point process on $\Rcal$ with intensity function
\begin{equation}\label{eq:def_intensity_function_f}
	f_{\alpha,\nu}(x,y) = \frac{\alpha \nu}{\pi} e^{-\alpha y}.
\end{equation}

We will write $p = (x, y)$ for points in $\Rcal$. Then the \emph{infinite limit model} $G_{\mathcal{P}}(\alpha, \nu)$ is defined to have vertex set $\mathcal{P}$ and edge set such that
\[
	(p_i, p_j) \in E(G_{\mathcal{P}}(\alpha, \nu)) \iff |x_i - x_j| \leq e^{\frac{y_i + y_j}{2}}.
\]
For a point $p \in \Rcal$, we will write $\BallPo{p}$ to denote the \emph{ball} around $p$, i.e.
\begin{equation}\label{eq:def_ball_P}
	\BallPo{p} = \{p^\prime \in \Rcal : |x - x^\prime| \leq e^{\frac{y + y^\prime}{2}}\}.
\end{equation}
With this notation we then have that $\BallPo{p} \cap \mathcal{P}$ denotes the set of neighbors of a vertex $p \in G_\Pcal(\alpha,\nu)$.  We denote the intensity measure of the Poisson process $\mathcal{P}$ by $\mu_{\alpha, \nu}$, i.e. for every Borel-measurable subset $S \subseteq \mathbb{R} \times \mathbb{R}_+$
\begin{equation}\label{eq:def_mu_P}
	\mu_{\alpha,\nu}(S) = \int_S f_{\alpha,\nu}(x,y) \, dx \, dy,
\end{equation}
and for any point $p \in \Rcal$ we let $\eta_p$ be the corresponding probability measure on $\BallPo{p}$ 
\begin{equation}\label{eq:def_etq_P}
	\eta_p(S) = \frac{\mu_{\alpha,\nu}(S \cap \BallPo{p})}{\mu_{\alpha,\nu}(\BallPo{p})}.
\end{equation}

Most of the subsequent analysis in this paper will make use of the~\emph{Campbell-Mecke} formula. 
For a Poisson point process $\mathcal{P}$ on a measurable space $S$ with intensity $\lambda$ and a measurable non-negative function $h: S^r \rightarrow \mathbb{R}$ we have
\begin{equation}\label{eq:def_Campbell-Mecke}
\begin{split}
& \Exp{\sum_{x_1,\ldots, x_r \in \mathcal{P}}^{\neq} h (x_1, \ldots, x_r,
\mathcal{P} \setminus \{x_1,\ldots, x_r\}) }\\
& = \int_S \cdots \int_S
\Exp{h(x_1,\ldots, x_r,\mathcal{P} \setminus \{x_1,\ldots, x_r\}) \lambda (x_1) \cdots \lambda (x_r) dx_1 \cdots dx_r},
\end{split}
\end{equation}
where the sum ranges over all distinct $r$-tuples of points of $\mathcal{P}$.

We shall use $\rho(p,k)$ and $\rho(y,k)$ (for a point $p=(x,y) \in \mathcal{R}$) to denote the probability mass function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(\BallPo{p})$, i.e.
\[
	\rho(p,k) = \rho(y,k) = \Prob{\Po\left(\xi_{\alpha,\nu} e^{y/2}\right) = k}.
\]
Note that $\rho(y,k)$ denote the probability that a node $(x,y) \in G_{\Pcal}(\alpha,\nu)$ has degree $k$.

\begin{remark}[Notations for points]\label{rmk:point_notation}
We will be working extensively with expressions in terms of points of $\Pcal_{\alpha, \nu}$, often relating them back to points in the hyperbolic disc $\Dcal_{\Rcal_n}$. Therefore, in the remainder of this paper we will always use $u = (r,\theta)$ to denote points in $\Dcal_{\Rcal_n}$ in polar coordinates while $p = (x,y)$ will denote points in $\Rcal := \mathbb{R} \times \mathbb{R}_+$ in Cartesian coordinates. In addition, when we write $p^\prime \in \mathbb{R} \times \mathbb{R}_+$ we will denote its Cartesian coordinates by $(x^\prime, y^\prime)$ and similarly for $p_i$, $u^\prime$ and $u_i$, e.g. $p_i = (x_i, y_i)$.
\end{remark}

\subsection{Finite box model}

Next we define the finite graph $G_{\mathcal{P},n}(\alpha, \nu)$ which we will couple to $G_{\H, n}(\alpha, \nu)$. Recall that $R_n = 2\log(n/\nu)$ and that $\Pcal_{\alpha, \nu}$ is a Poisson point process on $\Rcal = \R \times \R_+$, with intensity given by~\eqref{eq:def_intensity_function_f}. Now consider the finite box $\Rcal_n = (-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2}e^{R_n/2}] \times (0, R_n]$ in $\Rcal$. Then the \emph{finite box model} graph $G_{\mathcal{P},n}(\alpha, \nu)$ has vertex set $\mathcal{V}_n := \mathcal{P}_{\alpha, \nu} \cap \Rcal_n$ and edges set such that
\[
	(p_i, p_j) \in E(G_{\mathcal{P},n}(\alpha, \nu)) \iff |x_i - x_j|_{\pi e^{R_n/2}} \leq e^{\frac{y_i + y_j}{2}},
\]
where $|x|_{\pi e^{R_n/2}} = \inf_{k \in \mathbb{Z}} |x + k \pi e^{R_n/2}|$. In order words, $G_{\mathcal{P},n}(\alpha, \nu)$ is the subgraph of $G_{\mathcal{P}}(\alpha, \nu)$ induced on $\mathcal{V}_n$, where we have identified the boundaries of $(-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2}e^{R_n/2}]$ to exclude boundary effects and to make the model symmetric in the horizontal $x$-axis.

Similar to $G_{\Pcal}(\alpha,\nu)$, we define for any $p \in \Rcal_n$, its neighborhood ball
\[
	\BallPon{p} = \left\{p^\prime \in \Rcal_n : |x_i - x_j|_{\pi e^{R_n/2}} \leq e^{\frac{y_i + y_j}{2}}\right\}
\]
and associated functions $\mu_{\alpha,\nu ,n}$ and $\eta_{p,n}$ and others, by appending an additional subscript $n$. For instance
\[
	\mu_{\alpha,\nu,n}(S) = \Mu{S \cap \Rcal_n} \quad \text{and} \quad 
    \eta_{p,n}(S) = \frac{\mu_{\alpha,\nu ,n}(S \cap \BallPon{p}}{\mu_{\alpha,\nu,n}(\BallPon{p})}
    = \frac{\Mu{S\cap\BallPon{p}}}{\Mu{\BallPon{p}}}.
\]

@@

To couple the hyperbolic graph to $G_{\mathcal{P},n}(\alpha, \nu)$ we consider the original hyperbolic graph on a Poisson distributed number of vertices as an intermediate step. For this we define $G_{\HP,n}(\alpha, \nu)$ to be the hyperbolic random graph where the vertex set is given by $N \stackrel{d}{=} \Po(n)$ points, distributed according to the $(\alpha, R_n)$-quasi uniform distribution~\eqref{eq:def_hyperbolic_point_distribution} defined in Section \ref{ssec:hyperbolic_model}. The following two lemmas from \cite{fountoulakis2018law} establishes a coupling between this version of the hyperbolic random graph and the finite box model and relate the hyperbolic neighborhoods to the neighborhood balls $\BallPo{p}$. The ensuing proposition verifies that this coupling also essentially preserves the local clustering function, i.e. that the difference between the local clustering functions in the hyperbolic model and the finite box graph converges to zero faster than the proposed scalings in the main result, Theorem \ref{thm:asymptotic_local_clustering_hyperbolic}. Hence, it justifies the transition to the finite box model.

@@

These two results imply that the difference between the local clustering functions in the hyperbolic model and modified version in the finite box graph converges to zero faster than the proposed scaling $s_\alpha(k_n)$ in Theorem \ref{thm:asymptotic_local_clustering_hyperbolic}. Hence, to proof this theorem it enough to prove it for $c_{\mathcal{P},n}^\ast(k)$.

@@

With this concentration result, we are left with the task to compute the limit of $\Exp{c_{\mathcal{P},n}^\ast(k_n)}$ as $n \to \infty$ and show that it is equivalent to $c_\infty(k_n)$. To accomplish this we move to the infinite limit model $G_\Pcal(\alpha,\nu)$. The next result shows that the difference between the expected value of the clustering function $c_G^\ast(k)$ in $G_{\mathcal{P},n}(\alpha,\nu)$ and $G_{\mathcal{P}}(\alpha,\nu)$ goes to zero faster than the proposed scaling in Theorem \ref{thm:asymptotic_local_clustering_hyperbolic}.

\begin{proposition}[Transition to infinite limit model]\label{prop:convergence_average_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence tending to infinity and satisfying $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\infty(k_n)\right| = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

@@

We are now ready to prove our main results, using the propositions stated in the previous sections. We begin with Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}.

\begin{proof}[Proof of Theorem \ref{thm:asymptotic_local_clustering_hyperbolic}]
By applying Proposition~\ref{prop:convergence_average_clustering_P_n}, Proposition~\ref{prop:convergence_average_clustering_P_n} and Theorem~\ref{thm:asymptotics_average_clustering_P} we get
\begin{align*}
	&\hspace{-30pt}\Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}\\
	&\le \frac{\Exp{\left|c_{\Pcal,n}^\ast(k_n) - \Exp{}c_{\Pcal,n}^\ast(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)}
		+ \frac{\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\infty(k_n)\right|}{C_{\alpha,\nu}s_\alpha(k_n)}
		+ \left|\frac{c_{\infty}(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right| \\
	&= \smallO{1}.
\end{align*}
%\[
%	\Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}
%	\le \left|\frac{c_{\infty}(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right| 
%	+ \frac{\Exp{\left|c_{\Pcal,n}^\ast(k_n) - c_\infty(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)} = \smallO{1}.
%\]
Combining this with Proposition~\ref{prop:couling_c_H_P} yields
\[
	\Exp{\left|\frac{c_{\H,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}
	\le \Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|} 
	+ \frac{\Exp{\left|c_{\H,n}^\ast(k_n) - c_{\Pcal,n}^\ast(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)} = \smallO{1}.
\]
and the result then follows by applying Lemma~\ref{lem:clustering_ast_H}. 
\end{proof}

@@

Note that Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic} states that for any sequence $k_n  = \smallO{n^{1/(2\alpha + 1)}}$
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} = 0.
\]

Let $b_n := \log(n)^{-1} n^{1/(2\alpha + 1)}$ and define the sequence $k_n$ by
\[
	k_n = \arg \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}.
\]
Then $k_n = \smallO{n^{1/(2\alpha + 1)}}$ and by Theorem~\ref{thm:asymptotic_local_clustering_hyperbolic}
\[
	\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} = \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} \to 0.
\]
Similarly we note that
\[
	\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \to 0.
\]

>>>>

The key result in this paper is Theorem~\ref{thm:local_clustering_hyperbolic}. In particular, it is the key ingredient in the proofs of Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:asymptotics_average_clustering_P}. Therefore, in this section we first give a detailed overview of the structure for the proof of Theorem~\ref{thm:local_clustering_hyperbolic}. The proof itself, based on intermediate propositions, can be found in Section~\ref{ssec:proof_main_results}. Here we also prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. The proof of Theorem~\ref{thm:asymptotics_average_clustering_P} can be found in Section~\ref{sec:asymptotics_average_clustering_ast_P}.

Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:local_clustering_hyperbolic} based on the different propositions described below, plus the sections in which theses propositions are proved.

@@

\caption{Overview of the proof strategy for Theorem \ref{thm:local_clustering_hyperbolic}.}
\label{fig:overview_proof}
\end{figure}

The main idea behind the proof of Theorem \ref{thm:local_clustering_hyperbolic} is to couple the hyperbolic graph $G_{\H, n}(\alpha, \nu)$ with a finite subgraph $G_{\Pcal,n}(\alpha,\nu)$ of an infinite graph $G_\Pcal(\alpha,\nu)$ whose nodes are given by a Poisson Point Process (see Section~\ref{ssec:infinite_model}). The infinite graph will then yield the limit expressions for the local clustering coefficient and function. A similar approach is used in \cite{fountoulakis2018law} to analyze the size of the largest component in the hyperbolic model and in \cite{komjathy2018explosion} to analyze explosion times in a more general class of models.. 
The advantage of the infinite limit model is two-fold. First, it simplifies the actual computations, e.g. by replacing hyperbolic densities with exponential densities and the hyperbolic distance with a computationally easier expression. Secondly, it directly yields the limit as a result rather than having to determine this later from a long expression depending on $n$.

Very roughly, the proof falls into five parts: 1) replacing the randomly sampled nodes in the hyperbolic graph with a Poisson Point Process, 2) justifying the transitioning from this Poisson version of the hyperbolic random graph to the finite model, 3) proving concentration results for local clustering, 4) show that the limit expression correspond to those in the infinite model and finally 5) calculating the expected local clustering coefficient and function in the infinite limit model.

In this section, we state the required definitions, auxiliary lemmas and propositions and show how to assemble them to obtain Theorem~\ref{thm:local_clustering_hyperbolic}.

\subsection{Finite box model}\label{ssec:finite_model}

First we define the finite graph $G_{\mathcal{P},n}(\alpha, \nu)$ which we will couple to $G_{\H, n}(\alpha, \nu)$. As similar approach was used in~\cite{fountoulakis2018law} to analyze the largest component. Define $\Rcal = \mathbb{R} \times \mathbb{R}_+$ and let $\mathcal{P}_{\alpha,\nu}$ be a Poisson point process on $\Rcal$ with intensity function
\begin{equation}\label{eq:def_intensity_function_f}
	f_{\alpha,\nu}(x,y) = \frac{\alpha \nu}{\pi} e^{-\alpha y}.
\end{equation}

In addition we denote the intensity measure of the Poisson process $\mathcal{P}$ by $\mu_{\alpha, \nu}$, i.e. for every Borel-measurable subset $S \subseteq \mathbb{R} \times \mathbb{R}_+$
\begin{equation}\label{eq:def_mu_P}
	\mu_{\alpha,\nu}(S) = \int_S f_{\alpha,\nu}(x,y) \, dx \, dy,
\end{equation}
and for any point $p \in \Rcal$.

We will write $p = (x, y)$ for points in $\Rcal$. Then the \emph{infinite limit model} $G_{\mathcal{P}}(\alpha, \nu)$ is defined to have vertex set $\mathcal{P}$ and edge set such that
\[
	(p_i, p_j) \in E(G_{\mathcal{P}}(\alpha, \nu)) \iff |x_i - x_j| \leq e^{\frac{y_i + y_j}{2}}.
\]

\begin{remark}[Notations for points]\label{rmk:point_notation}
We will be working extensively with expressions in terms of points of $\Pcal_{\alpha, \nu}$, often relating them back to points in the hyperbolic disc $\Dcal_{\Rcal_n}$. Therefore, in the remainder of this paper we will always use $u = (r,\theta)$ to denote points in $\Dcal_{\Rcal_n}$ in polar coordinates while $p = (x,y)$ will denote points in $\Rcal := \mathbb{R} \times \mathbb{R}_+$ in Cartesian coordinates. In addition, when we write $p^\prime \in \mathbb{R} \times \mathbb{R}_+$ we will denote its Cartesian coordinates by $(x^\prime, y^\prime)$ and similarly for $p_i$, $u^\prime$ and $u_i$, e.g. $p_i = (x_i, y_i)$. In addition, for a point $p = (x,y)$, we will refer to $y$ as the \emph{height} of $p$.
\end{remark}

For the definition of the finite graph, recall that $R_n = 2\log(n/\nu)$ and consider the finite box $\Rcal_n = (-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2}e^{R_n/2}] \times (0, R_n]$ in $\Rcal$. Then the \emph{finite box model} graph $G_{\mathcal{P},n}(\alpha, \nu)$ has vertex set $\mathcal{V}_n := \mathcal{P}_{\alpha, \nu} \cap \Rcal_n$ and edges set such that
\[
	(p_i, p_j) \in E(G_{\mathcal{P},n}(\alpha, \nu)) \iff |x_i - x_j|_{\pi e^{R_n/2}} \leq e^{\frac{y_i + y_j}{2}},
\]
where $|x|_{\pi e^{R_n/2}} = \inf_{k \in \mathbb{Z}} |x + k \pi e^{R_n/2}|$. In order words, $G_{\mathcal{P},n}(\alpha, \nu)$ is the subgraph of $G_{\mathcal{P}}(\alpha, \nu)$ induced on $\mathcal{V}_n$, where we have identified the boundaries of $(-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2}e^{R_n/2}]$ to exclude boundary effects and to make the model symmetric in the horizontal $x$-axis.

For a point $p \in \Rcal_n$, we will write $\BallPon{p}$ to denote the \emph{ball} around $p$, i.e.
\begin{equation}\label{eq:def_ball_P_n}
	\BallPon{p} = \left\{p^\prime \in \Rcal_n : |x_i - x_j|_{\pi e^{R_n/2}} \leq e^{\frac{y_i + y_j}{2}}\right\}.
\end{equation}
With this notation we then have that $\BallPon{p} \cap \PPP$ denotes the set of neighbors of a vertex $p \in G_{\Pcal,n}(\alpha,\nu)$.

@@

To couple the hyperbolic graph to $G_{\mathcal{P},n}(\alpha, \nu)$ we consider the original hyperbolic graph on a Poisson distributed number of vertices as an intermediate step. For this we define $G_{\HP,n}(\alpha, \nu)$ to be the hyperbolic random graph where the vertex set is given by $N \stackrel{d}{=} \Po(n)$ points, distributed according to the $(\alpha, R_n)$-quasi uniform distribution~\eqref{eq:def_hyperbolic_point_distribution} defined in Section \ref{ssec:hyperbolic_model}. The following two lemmas from \cite{fountoulakis2018law} establishes a coupling between this version of the hyperbolic random graph and the finite box model and relate the hyperbolic neighborhoods to the neighborhood balls $\BallPo{p}$. The ensuing proposition verifies that this coupling also essentially preserves the local clustering function, i.e. that the difference between the local clustering functions in the hyperbolic model and the finite box graph converges to zero faster than the proposed scalings in the main result, Theorem \ref{thm:local_clustering_hyperbolic}. Hence, it justifies the transition to the finite box model.

@@

These two results imply that the difference between the local clustering functions in the hyperbolic model and modified version in the finite box graph converges to zero faster than the proposed scaling $s_\alpha(k_n)$ in Theorem \ref{thm:local_clustering_hyperbolic}. Hence, to proof this theorem it enough to prove it for $c_{\mathcal{P},n}^\ast(k)$.

To analyze the limiting expressions for clustering in the finite random graph $G_{\Pcal,n}(\alpha, \nu)$ we shall use an infinite graph model which can be understood as taking the finite graph $G_{\Pcal,n}(\alpha,\nu)$ and taking $n \to \infty$. Recall that $\Rcal = \mathbb{R} \times \mathbb{R}_+$ and that $\mathcal{P}_{\alpha,\nu}$ is a Poisson point process on $\Rcal$ with intensity function
\[
	f_{\alpha,\nu}(x,y) = \frac{\alpha \nu}{\pi} e^{-\alpha y}.
\]

The \emph{infinite limit model} $G_{\mathcal{P}}(\alpha, \nu)$ is defined to have vertex set $\mathcal{P}$ and edge set such that
\[
	(p_i, p_j) \in E(G_{\mathcal{P}}(\alpha, \nu)) \iff |x_i - x_j| \leq e^{\frac{y_i + y_j}{2}}.
\]
Similar to the finite graph, we write $\BallPo{p}$ to denote the \emph{ball} around $p$, i.e.
\begin{equation}\label{eq:def_ball_P}
	\BallPo{p} = \{p^\prime \in \Rcal : |x - x^\prime| \leq e^{\frac{y + y^\prime}{2}}\}.
\end{equation}

To analyze expression in the both the finite and infinite model we will make use of the~\emph{Campbell-Mecke} formula. 
For a Poisson point process $\mathcal{P}$ on a measurable space $S$ with intensity $\lambda$ and a measurable non-negative function $h: S^r \rightarrow \mathbb{R}$ we have
\begin{equation}\label{eq:def_Campbell-Mecke}
\begin{split}
& \Exp{\sum_{x_1,\ldots, x_r \in \mathcal{P}}^{\neq} h (x_1, \ldots, x_r,
\mathcal{P} \setminus \{x_1,\ldots, x_r\}) }\\
& = \int_S \cdots \int_S
\Exp{h(x_1,\ldots, x_r,\mathcal{P} \setminus \{x_1,\ldots, x_r\}) \lambda (x_1) \cdots \lambda (x_r) dx_1 \cdots dx_r},
\end{split}
\end{equation}
where the sum ranges over all distinct $r$-tuples of points of $\mathcal{P}$.

We shall use $\rho(p,k)$ and $\rho(y,k)$ (for a point $p=(x,y) \in \mathcal{R}$) to denote the probability mass function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(\BallPo{p})$, i.e.
\[
	\rho(p,k) = \rho(y,k) = \Prob{\Po\left(\xi_{\alpha,\nu} e^{y/2}\right) = k}.
\]
Note that $\rho(y,k)$ denote the probability that a node $(x,y) \in G_{\Pcal}(\alpha,\nu)$ has degree $k$, which due to the translation symmetry of the Poisson Point Process in the $x$ direction depends only on the \emph{height} $y$.

\begin{remark}[Degree distributions in the other models]
Similar to the infinite graph, the degree of a node $p$ in the Poisson hyperbolic graph $G_{\HP,n}$ and the finite random graph $G_{\Pcal,n}$ is distributed as a Poisson random variable with mean $\Mu{\BallHyp{p}}$ and $\Mu{\BallPon{p}}$, respectively. Following the notation above we will denote these distributions by $\rho_{\HP,n}(y,k)$ and $\rho_{\Pcal,n}(y,k)$, since these again are independent of the $x$-coordinate of $p$.
\end{remark}

We can also define a local clustering coefficient and function in this model. However, since $G_\Pcal$ has an infinite number of nodes the corresponding definitions are slightly different than for the finite size models. Intuitively the local clustering coefficient is the probability that two random neighbors of a randomly sampled node form a triangle, while the local clustering function is the same probability but conditioned on the sampled node having degree $k$.

Again, due to symmetry in the $x$ direction we can assume without loss of generality that a randomly sampled point $p$ is $(0,y)$. For any $y \in \R_+$ we define the probability density function on $\Rcal := \R_+ \times \R$ by  
\begin{equation}\label{eq:def_eta_P}
	\eta_y(x^\prime, y^\prime) = \frac{\ind{p^\prime \in \BallPo{(0,y)}} f_{\alpha,\nu}(x^\prime,y^\prime)}{\mu_{\alpha,\nu}(\BallPo{(0,y)})}.
\end{equation}
That is $\eta_y$ is the density $f_{\alpha,\nu}$ restricted to the ball $\BallPo{(0,y)}$, properly normalized. Or in other words, it denotes the probability density of a randomly sampled point in $\BallPo{(0,y)}$, i.e. a randomly sampled neighbor of $(0,y)$.

We then define
\begin{equation}\label{eq:def_delta_p}
	\Delta_{\Pcal}(y) = \iint_{\Rcal^2} T_\Pcal(y, x_1, x_2, y_1, y_2) \eta_{y}(x_1,y_1)\eta_{y}(x_2,y_2) \dd x_1 \dd x_2  \dd y_1  \dd y_2,
\end{equation}
with 
\[
	T_\Pcal(y, x_1, x_2, y_1, y_2) 
	= \ind{|x_1| \, \le \, e^{(y+y_1)/2}}\ind{|x_2| \, \le \, e^{(y+y_2)/2}}\ind{|x_1-x_2| \, \le \, e^{(y_1+y_2)/2}},
\]
which is the probability that two randomly sampled points in $\BallPo{(0,y)}$ form a triangle with $(0,y)$.

Finally, we note that the height $y$ of a randomly sampled point has the exponential density $\alpha e^{-\alpha y}$. \PvdH{Could someone add an intuition or better explanation why this is the case.}

With these expressions we define the limiting local clustering coefficient as
\begin{equation}\label{eq:def_average_clustering_infinite_model}
	c_\Pcal := \int_0^\infty \Delta_{\Pcal}(y) (1 - \rho(y,0) - \rho(y,1)) \alpha e^{-\alpha y} \dd y,
\end{equation}
while the limiting local clustering function is defined as
\begin{equation}\label{eq:def_c_infty_k}
	c_\Pcal(k) = \frac{\int_0^\infty \rho(y,k) \Delta_\Pcal(y) \alpha e^{-\alpha y} \dd y}{\int_0^\infty \rho(y,k) \alpha e^{-\alpha y} \dd y}.
\end{equation}

In Section~\ref{sec:asymptotics_average_clustering_ast_P} we will compute both these quantities and show that $c_\Pcal = c_\infty$ and $c_\Pcal(k) = c_\infty(k)$, where $c_\infty$ and $c_\infty(k)$ are given in, respectively, Theorem~\ref{thm:clustering_coefficient_hyperbolic} and Theorem~\ref{thm:local_clustering_hyperbolic}

@@

With this concentration result, we are left with the task to compute the limit of $\Exp{c_{\mathcal{P},n}^\ast(k_n)}$ as $n \to \infty$ and show that it is equivalent to $c_\infty(k_n)$. To accomplish this we move to the infinite limit model $G_\Pcal(\alpha,\nu)$ and show that the difference between the expected value of the clustering function $c_G^\ast(k)$ in $G_{\mathcal{P},n}(\alpha,\nu)$ and $G_{\mathcal{P}}(\alpha,\nu)$ goes to zero faster than the proposed scaling in Theorem \ref{thm:local_clustering_hyperbolic}.

\begin{proposition}[Transition to infinite limit model]\label{prop:convergence_average_clustering_P_n}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and let $(k_n)_{n \ge 1}$ be any positive sequence tending to infinity and satisfying $k_n = \bigO{n^{\frac{1}{2\alpha+1}}}$. Then, as $n \to \infty$,
\[
	\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\Pcal(k_n)\right| = \smallO{s_\alpha(k_n)}.
\]
\end{proposition}

@@

We are now ready to prove our main results, using the propositions stated in the previous sections. We begin with Theorem~\ref{thm:local_clustering_hyperbolic}.

\begin{proof}[Proof of Theorem \ref{thm:local_clustering_hyperbolic}]
By applying Proposition~\ref{prop:convergence_average_clustering_P_n}, Proposition~\ref{prop:convergence_average_clustering_P_n} and Theorem~\ref{thm:asymptotics_average_clustering_P} we get
\begin{align*}
	&\hspace{-30pt}\Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}\\
	&\le \frac{\Exp{\left|c_{\Pcal,n}^\ast(k_n) - \Exp{}c_{\Pcal,n}^\ast(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)}
		+ \frac{\left|\Exp{c_{\Pcal,n}^\ast(k_n)} - c_\infty(k_n)\right|}{C_{\alpha,\nu}s_\alpha(k_n)}
		+ \left|\frac{c_{\infty}(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right| \\
	&= \smallO{1}.
\end{align*}
%\[
%	\Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}
%	\le \left|\frac{c_{\infty}(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right| 
%	+ \frac{\Exp{\left|c_{\Pcal,n}^\ast(k_n) - c_\infty(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)} = \smallO{1}.
%\]
Combining this with Proposition~\ref{prop:couling_c_H_P} yields
\[
	\Exp{\left|\frac{c_{\H,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|}
	\le \Exp{\left|\frac{c_{\Pcal,n}^\ast(k_n)}{C_{\alpha,\nu}s_\alpha(k_n)} - 1\right|} 
	+ \frac{\Exp{\left|c_{\H,n}^\ast(k_n) - c_{\Pcal,n}^\ast(k_n)\right|}}{C_{\alpha,\nu}s_\alpha(k_n)} = \smallO{1}.
\]
and the result then follows by applying Lemma~\ref{lem:clustering_ast_H}. 
\end{proof}

@@

Note that Theorem~\ref{thm:local_clustering_hyperbolic} states that for any sequence $k_n  = \smallO{n^{1/(2\alpha + 1)}}$
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} = 0.
\]

Let $b_n := \log(n)^{-1} n^{1/(2\alpha + 1)}$ and define the sequence $k_n$ by
\[
	k_n = \arg \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}.
\]
Then $k_n = \smallO{n^{1/(2\alpha + 1)}}$ and by Theorem~\ref{thm:local_clustering_hyperbolic}
\[
	\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} = \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} \to 0.
\]
Similarly we note that
\[
	\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \to 0.
\]

Now define
\[
	a_n := \left\lfloor \min \left\{
	\left(\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}\right)^{-1/2},
	\left(\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|}\right)^{-1/2},
	b_n - 1 \right\} \right\rfloor
\]
so that $a_n < b_n = \smallO{n^{1/(2\alpha + 1)}}$, $a_n \to \infty$, 
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} 
	\le a_n \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} \to 0,
\]
and similarly
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \to 0.
\]
